---
title: "Reproducible Research Week 2 Course Project 1"
author: "Mohammad Umer Ayub"
date: "October 9, 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading and preprocessing the data
##1. Load the data
```{r}
x_data<-read.csv(file ="activity.csv")
```

##2. Process/transform the data
```{r}
x_data$date= as.Date(x_data$date, "%Y-%m-%d")
summary(x_data)
```

##What is mean total number of steps taken per day
##1. Calculate the total number of steps taken per day
```{r}
X_data_agg_date<-aggregate(x_data$steps,by=list(x_data$date),FUN=sum, na.rm= TRUE )
X_data_agg_date
```

##2. Make a histogram of the total number of steps taken each day
```{r}
hist(X_data_agg_date$x,main="Histogram:Total number of steps per day", 
     xlab="Total Number of steps in a day")
```

##3. Calculating mean and Median
```{r}
##mean
mean(X_data_agg_date$x)
##Median
median(X_data_agg_date$x)
```


##What is the average daily activity pattern?
##1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
##The average number of steps taken, averaged across all days
X_data_agg_intv<-aggregate(x_data$steps,by=list(x_data$interval),FUN=mean, na.rm= TRUE )
##Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and  (y-axis) 
plot(X_data_agg_intv$Group.1,X_data_agg_intv$x,type = 'l',
main="Average number of steps over all days", xlab="5 Min Interval", 
ylab="Average number of steps")
```

##2. On average across all the days in the dataset, contains the maximum number of steps
```{r}
X_data_agg_intv[which.max(X_data_agg_intv$x),]
```


##Imputing missing values
##1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(x_data$steps))
```

##2. Devise a strategy for filling in all of the missing values in the dataset.
##Strategy : Replacing the NA with MEAN of that specific Interval

##3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
x_data_new<-x_data
for (i in 1:nrow(x_data_new)) {
        if (is.na(x_data_new$steps[i])) {
                
                intval_val <- x_data_new$interval[i]
                steps_val <- X_data_agg_intv[X_data_agg_intv$Group.1 == intval_val,]
                x_data_new$steps[i] <- steps_val$x
        }
}
```

##4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
##Taking aggregate day by day
X_data_agg_date_new<-aggregate(x_data_new$steps,by=list(x_data_new$date),FUN=sum, na.rm= TRUE )
##Make a histogram of the total number of steps taken each day (Imputed)
hist(X_data_agg_date_new$x,main="Histogram:Total number of steps per day (Imputed)", 
     xlab="Total Number of steps in a day")
##Calculating mean
mean(X_data_agg_date_new$x)
##Calculating median
median(X_data_agg_date_new$x)
##Yes the mean and median are have changed.
```

##Are there differences in activity patterns between weekdays and weekends?
##1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r}
x_data_new$day <- weekdays(x_data$date)
x_data_new$day[x_data_new$day%in% c('Saturday','Sunday') ] <- "Weekend"
x_data_new$day[x_data_new$day != "Weekend" ] <- "Weekday"
x_data_new$day_type <- as.factor(x_data_new$day)
```

##2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
X_data_agg_new_weekend_fact <- aggregate(steps ~ interval + day_type, x_data_new, mean)
library(ggplot2)
qplot(interval,steps,data = X_data_agg_new_weekend_fact, type = 'l',  geom=c("line"),xlab = "Interval", 
      ylab = "Number of steps", main = "") + facet_wrap(~ day_type, ncol = 1)
```
##Yes, Activity paterns are different Weekdays and Weekends